\large{\bf Large effects and the infinitesimal model }

### 2023-07-07 writeup.tex
FIxed typo
minor edits up to section 4

The Gaussian distribution appears thanks to ``the'' Central Limit Theorem,
i.e., the universality of adding up a great many very small and independent things.
However, there is a wider class of Central Limit Theorems,
which describe the distributions of sums of a great many independent and exchangeable things:
the \emph{stable laws} \citep{gnedenko1968limit}
Roughly speaking, these say that if we add together $n$ independent copies of a random variable $X$
for which $\P\{ X > t \}$ is proportional to $t^{-\alpha}$ for some $0 < \alpha < 2$,
then -- almost regardless of the details -- the result, scaled by $n^{1/\alpha}$,
is well-approximated by a universal form, an $\alpha$-stable distribution
(and the approximation becomes exact, as $n \to \infty$).
In such ensembles, single large entries can be important:
concretely, the largest of the $n$ copies will be of order $n^{1/\alpha}$,
so that e.g., for $\alpha=1$ a positive proportion of the sum of $n$ entries will be contributed by a single one.

@@

Stable distributions are by no means new to population biology.
For instance, it is a long-standing question how often the motions of organisms
are better modeled by essentially Brownian random walks or by L\'evy processes \citep{benhamou2007animals}
and a growing literature studies the effects of such power law distributions
on spatial genetic and dynamics \comment{I'm not quite sure what you mean by that}
\citep[e.g.,][]{paulose2019spatial,smith2023isolation}.
More directly analogous to the problem we study here, \citet{landis2012phylogenetic}
found that an $\alpha$-stable process (with $\alpha \approx 1.5$) fit the evolution of log endocranial volume
on the primate phylogeny
better than either a Brownian model.
It is also important to note that
the infinitesimal model implies a Gaussian distribution for the difference between an offspring's trait and their parents' average,
\emph{not} for the entire population.
The question of when the trait distribution in an evolving population
might be modeled as a Gaussian has been discussed extensively
\citep[e.g.,][]{turelli1994genetic,dbarre2015evolution}.

The ways in which genetic variants affect an organisms' trait can be complex,
and interactions between alleles, between loci, and with the environment are often important in practice.
In fact, many traits claimed to be genetically determined
in fact only appear heritable through non-genetic means,
e.g., systematic inequalities in resource distribution partitioned by class or race.
% \citep[e.g.,][]{davenport1916hereditary}.
In this paper, we follow the literature in working with the relatively simple purely additive case
of a trait that can be decomposed into a sum across loci of the contributions of the alleles from each locus,
even omitting an independent effect of ``the environment''.
However, we emphasize that such models should \emph{not} always be the default for understanding biological variation,
especially for human traits.

@@

Suppose that the focal individual's trait is
$$
    Z_M = \sum_{\ell=1}^M \eta_\ell / M^{1/\alpha},
$$
where $\eta_\ell$ is the contribution of the allele carried at the $\ell^\text{th}$ locus.
The effects of the alleles carried by the two parents
are $\eta_\ell^{[1]}$ and $\eta_\ell^{[2]}$, respectively,
and their traits are
$Z_M^{[1]} = \sum_\ell \eta_\ell^{[1]} / M^{1/\alpha}$ and
$Z_M^{[2]} = \sum_\ell \eta_\ell^{[2]} / M^{1/\alpha}$.
We assume the loci are unlinked and inheritance is fair, so
letting $X_\ell$ be a family of Bernoulli(1/2) random variables that are independent of everything else,
$$
    \eta_\ell = X_\ell \eta_\ell^{[1]} + (1 - X_\ell) \eta_\ell^{[2]} .
$$
What we will call the \emph{transmission noise}, $R_M$,
is the offspring's trait minus the midparent value, $\bar Z_M$:
\begin{align*}
    \bar Z_M &= \frac{Z_M^{[1]} + Z_M^{[2]}}{2} \\
    R_M &= Z_M - \bar Z_M \\
        &= \sum_{\ell=1}^M
            \frac{
                X_\ell (\eta_\ell^{[1]} - \eta_\ell^{[2]})
                + (1 - X_\ell) (\eta_\ell^{[2]} - \eta_\ell^{[1]})
            }{ 2 } .
\end{align*}

@@

It is mathematically intriguing to speculate what effect a power law effect size distribution
for would have on the induced evolutionary dynamics.
But, is there any evidence that non-Gaussian transmission noise is important in practice?
In this paper we focus primarily on the theory,
but first take a brief and imperfect look at what some data have to say on the topic.
To do this, we downloaded the GWAS results provided by~\citet{biobankSNPs}
for 7,221 phenotypes from 361,194 human individuals from the March 2018 release of the UK Biobank.
After filtering (see details below) we were left with 695 binary illness-related phenotypes.
Taken at face value, these provide for each phenotype a set of estimates of the additive effect
of the alternative allele at a set of single nucleotide polymorphisms (SNPs)
on the phenotype relative to the reference allele.
The phenotypes we consider were all binary and coded as 0/1,
so estimates could be thought of as additive adjustments to the risk
(so, effects are \emph{not} on a logit scale).
These estimates were obtained as the coefficient for the SNP in a simple linear model
fit to each phenotype for each SNP,
with the first 20 principal components and all combinations of age, age squared, and inferred sex as covariates;
we considered effects of those SNPs with a reported $p$-value less than $10^{-8}$.
There are a great many possible issues with these data;
however, the results seem consistent across many SNPs and phenotypes,
and we know of no reason that possible methodological artifacts
(e.g., confounding by uncorrected variables)
would specifically affect the decay of the tail of effect sizes.

Since the transmission noise is a sum of terms of the form
$$
    (X_\ell - 1/2) \eta_\ell^{[1]} + (1/2 - X_\ell) \eta_\ell^{[2]},
$$
where $\eta_\ell^{[i]}$ is the effect of the allele in parent $i$ at locus $\ell$,
the appropriate central limit theorem for the sum is determined by
the tail behavior of $|(X - 1/2) \eta| = |\eta|/2$,
where $X$ is Bernoulli(1/2) and $\eta$ is the effect
of a uniformly chosen allele at a uniformly chosen locus.
So, we'd like to know
if $\P\{|X|>2t\}$ is proportional to $t^{-\alpha}$,
i.e., the value of
$$
    \lim_{t \to \infty} \frac{ \log \P\{ |X| > 2 t \} }{ \log t } = \alpha ,
$$
if the limit exists.
Let $p_\ell$ be the minor allele frequency at the $\ell^\text{th}$ locus;
since the data we have assigns effect 0 to the major allele,
$\eta_\ell = 0$ with probability $1-p_\ell$.
Therefore, if we have estimated effects for $M$ SNPs
with minor allele frequency $p_\ell$ and effect size $e_\ell$ for the minor allele,
for $1 \le \ell \le M$,
then
$$
    \P\{ |\eta| > 2t \} \approx \frac{1}{M} \sum_{\ell : |e_\ell| > 2t} p_\ell .
$$
So, to obtain a per-trait estimate of $\alpha$,
we plot $\log \frac{1}{M} \sum_{\ell : |e_\ell| > t} p_\ell$ against $\log t$
for a sequence of values of $t$,
and fit a linear model to the tail
(which we take to be all values of $t$ for which the number of SNPs with absolute effect above $t$
is between 20 and 10\% of all SNPs).
Examples are shown in Figure~\ref{fig:example_snps}.

@@

Perhaps surprisingly, there is a converse --
in at least some situations,
independence of transmission noise implies a Gaussian distribution.
To see this,
divide the parental loci into groups according to the results of inheritance,
defining
\begin{align*}
    U_M^{[1]} &= \sum_{\ell=1}^M X_\ell \eta_\ell^{[1]}  ,
    \qquad \qquad
    V_M^{[1]} = \sum_{\ell=1}^M (1-X_\ell) \eta_\ell^{[1]} ,  \\
    U_M^{[2]} &= \sum_{\ell=1}^M (1-X_\ell) \eta_\ell^{[2]} ,
    \qquad \qquad
    V_M^{[2]} = \sum_{\ell=1}^M X_\ell \eta_\ell^{[2]}  ,
\end{align*}
so that 
$Z_M^{[1]} = U_M^{[1]} + V_M^{[1]}$ and
$Z_M^{[2]} = U_M^{[2]} + V_M^{[2]}$, while
$Z_M = U_M^{[1]} + U_M^{[2]}$ is the transmitted portion of the parental genomes,
and we can define $V_M = V_M^{[1]} + V_M^{[2]}$ to be the untransmitted portion.
Furthermore,
\begin{align*}
    \bar Z_M = \frac{Z_M + V_M}{2}
    \qquad \text{ and } \qquad
    R_M = \frac{Z_M - V_M}{ 2 } .
\end{align*}

@@

The Kac-Bernstein theorem states that if two independent random variables
have the property that if their sum and difference are independent of each other,
then the random variables have a joint Gaussian distribution
(\citet{kac1939characterization,bernstein1941property},
and see \citet{kagan1973characterization} for many much more general results).
Since $R_M$ is (half) the difference and $\bar Z_M$ is (half) the sum
of $Z_M$ and $V_M$, and $Z_M$ and $V_M$ are independent,
if we assume that also $R_M$ and $\bar Z_M$ are independent,
then both are therefore jointly Gaussian.

@@

\begin{prop}\label{prop:parental_contribs}
    Suppose that $\{\eta_\ell^{[i]}\}$ are independent for all $\ell$ and $i$,
    that the distribution of $\eta_\ell^{[i]}$ only depends on $\ell$,
    and that $\{X_\ell\}$ are independent of $\{\eta_\ell^{[i]}\}$.
    If the transmission noise $R_M$ is independent of the midparent value $\bar Z_M$,
    then $R_M$ and $\bar Z_M$ are jointly Gaussian.
\end{prop}

@@

Since models can be easily set up for which trait distributions are not Gaussian,
the implication of this is that for such models, independence of transmission noise
is not likely a good assumption.
Note that the proof does not depend on independence of the $X_\ell$:
indeed, alleles may be inherited together,
although in practice this would quickly create dependencies between nearby allelic effects.
Also note that it is \emph{not} usually true that the components --
for instance, $U_M^{[1]}$ and $U_M^{[2]}$ --
are independent of each other.

The presence of a large effect polymorphic allele
makes $R_M$ and $\bar Z_M$ dependent:
consider the extreme case where $\eta^{[i]}_\ell$ are i.i.d.{} with variance 1 for $\ell \ge 2$,
while $\eta^{[i]}_1 = \pm 10$ with probability 1/2 each,
and $Z^{[i]} = \eta_1^{[i]} + \sum_{\ell=2}^M \eta_\ell^{[i]} / \sqrt{M}$.
Then in the limit $R_M$ depends strongly on parental trait values:
if $Z^{[1]} \approx Z^{[2]}$ and so $\bar Z \approx \pm 10$ then most likely
$\eta_1^{[1]} = \eta_1^{[2]}$ and so $R$ is Normal with variance 1/2;
but if $\eta_1^{[1]} \neq \eta_1^{[2]}$ (and so $\bar Z \approx 0$)
then $R$ is close to $+5$ or $-5$ with equal probability.

@@

Concretely, consider a  Moran process where an individual whose trait value is $Z$ dies at rate $\mu(Z)$
and is replaced by a the offspring of two randomly chosen individuals,
i.e., a new individual of phenotype
\begin{align*}
  \frac{Z^{1} +Z^{2}}{2} + R ,
\end{align*}
where $Z^{1}$ and $Z^{2}$ are independent draws from the empirical population distribution
and $R$ is an independent draw from the Mendelian sampling distribution
(independent of $Z^{1}$ and $Z^{2}$). 
Mutation-selection balance demands that that
the distribution of individuals removed by death
and the distribution of individuals added by birth are equal,
\ie that $\frac{Z^{1} +Z^{2}}{2} + R$ is equal in distribution
to a $\mu$-weighted choice of $Z$.
% In other words, for all bounded and continuous functions $f$,
% \begin{equation} \label{eqn:MSB_condition}
%     \E\left[ f\left(\frac{Z^{1} +Z^{2}}{2} + R \right) \right] = \frac{\E\left[ f(Z) \mu(Z) \right]}{\E[\mu(Z)]},
% \end{equation}
% where the expectation is with respect to the joint law of the population empirical distribution (the law of $Z$, $Z^{1}$ and $Z^{2}$), and $R$.
% 
% Since the set of functions $\mathcal{S} = \{e^{i \theta x} : \theta \in \R\}$ is separating (see \eg \S 3.4 in \cite{Ethier+Kurtz1986}), it suffices to show that \eqref{eqn:MSB_condition} holds for $f \in \mathcal{S}$, \ie
Taking the characteristic function of both these distributions,
deterministic mutation-selection balance is therefore equivalent to
\begin{align} \label{eqn:MSB_CF}
    \E\left[ e^{i\theta\left(\frac{Z^{1} +Z^{2}}{2} + R \right)}\right] = \frac{\E\left[ e^{i \theta Z}\mu(Z) \right]}{\E[\mu(Z)]}.
\end{align}
Now, we assumed that $Z^{1}$, $Z^{2}$, and $R$ are independent, so we can write \eqref{eqn:MSB_CF} as 
\begin{align*}
    \frac{\E\left[ e^{i\theta Z} \mu(Z) \right]}{\E[\mu(Z)]} &= \E\left[ e^{i\theta\left(\frac{Z^{1} +Z^{2}}{2} + R \right)}\right]\\
	% &= \E\left[ e^{i\frac{\theta}{2} Z^{1}}e^{i\frac{\theta}{2} Z^{2}}e^{\i\theta R}\right]\\
	&= \E\left[e^{i\frac{\theta}{2} Z^{1}}\right]\E\left[e^{i\frac{\theta}{2} Z^{2}}\right]\E\left[e^{i\theta R}\right]\\
	&= \phi\left(\frac{\theta}{2}\right)^{2} \Phi(\theta),
\end{align*}
where
$ \phi(\theta) = \mathbb{E}\left[e^{i\theta Z^{1}}\right] = \mathbb{E}\left[e^{i\theta Z^{2}}\right] $
and
$ \Phi(\theta) = \mathbb{E}\left[e^{i\theta R}\right] $
are the characteristic functions of $Z^{j}$ and $R$, respectively.  In what follows, we will explore some non-Gaussian solutions $\phi(\theta)$ to \eqref{eqn:MSB_CF}.

@@

The key observation is this:
while is meaningless to speak of variances when the transmission noise is $\alpha$-stable, if an individual's trait value is informative of the allelic effect at any given locus,
then its trait value is informative of the \emph{distribution} of an offsprings' trait value,
and hence there can be no independence between parental traits and the Mendelian sampling term
that is central to the tractability of the infinitesimal model.

@@

\begin{figure}
    \begin{center}
        \includegraphics{sims/neutral_seg_noise}
    \end{center}
    \caption{
        Segregation noise with Gaussian \textbf{(A-C)} and Cauchy \textbf{(D-F)} effect size distributions,
        from the same simulations shown in Figure~\ref{fig:trait_distrns},
        for all individuals born during the last 1,000 time units of the simulation.
        \textbf{(A,D)} Segregation noise (offspring trait minus midparent) plotted against midparent.
        \textbf{(B,E)} Histograms of transmission noise.
        \textbf{(B,E)} Enrichment in quantiles of transmission noise conditional on midparent value.
        To compute the latter,
        let $q_0, q_2, q_4, \ldots, q_{98}, q_{100}$ divide the distribution of transmission noise
        into 50 bins with roughly equal numbers in each (i.e., the quantiles),
        let $p_k(a,b)$ denote the proportion of the offspring of parents with midparent value
        is in $[a, b)$ whose transmission noise is in $[q_{k-1}, q_k)$,
        and let $p_k$ be the proportion of all offspring with segregation noise in $[q_{k-1}, q_k)$
        (so, $p_k \approx 0.02$).
        Then, each line shows $p_k(a,b)/p_k$
        plotted against the midpoint of the corresponding quantile bin $(q_{k-1},q_k)$,
        for $(a,b)$ chosen to span the 5\% of midparents centered on:
        (red) the 10\% quantile,
        (black) the median, and
        (green) the 90\% quantile of midparent value.
        \label{fig:seg_noise}
    }
\end{figure}

We first consider the neutral case (i.e., with $\mu \equiv 1$).
As shown in Figure~\ref{fig:trait_distrns},
a simulation with a Cauchy effect size distribution, unsurprisingly,
has a distribution of trait values with more extreme values
and a rougher path of population median trait value over time
than does a simulation with a Gaussian effect size distribution.
Indeed, the sum of all effects along the path from the start of the simulation
to an arbitrary individual will approximate either Brownian motion (in the Gaussian case)
or a Cauchy process (in the Cauchy case).
Figure~\ref{fig:seg_noise} shows how transmission noise
depends on midparent value.
In the Gaussian case, transmission noise is Gaussian and independent of midparent value,
as shown by the horizontal lines at 1.0 in Figure~\ref{fig:seg_noise}C. 
On the other hand, in the Cauchy case we can see from either Figure~\ref{fig:seg_noise}D or F
that transmission noise depends strongly on midparent trait,
as we would expect in the presence of several large-effect alleles.

@@

However,
the two models look much more similar when considering only alleles of small effect.
Figure~\ref{fig:seg_noise_small} shows that ``midparent value''
and ``transmission noise'' computed using only mutations with effect size less than 0.2
(excluding less than 1\% of mutations)
are quite close to independent.
The result does not depend strongly on the cutoff chosen,
and are similar even if less 0.1\% of mutations are excluded.

@@

Further analysis of the infinitesimal model with other noise distributions,
as we have begun here,
could be interesting,
e.g., to see how changing the distribution of transmission noise
affects the rate of adaptive evolution,
levels of genetic load,
the rate of fixation,
and the strength of linked selection.
However,
we have seen that the assumption of the infinitesimal model
implies a sort of ``blending inheritance''
for the large effect alleles,
and so such work must be accompanied by simulations
to see if the predictions are borne out under concrete models of genetic inheritance.
Another open question is whether there is some set of assumptions
(perhaps involving linkage, pleiotropy, and/or environmental interactions with genetic effects)
which leads to these non-Gaussian infinitesimal models.
Finally, it may well be that there is a clever way to set up a trait-only model of evolution
that is both insensitive to the underlying genetics
and does not rely on the Gaussian central limit theorem.

>>>>

The Gaussian distribution appears thanks to ``the'' Central Limit Theorem,
i.e., the universality of adding up a great many very small and independent things.
However, there is a wider class of Central Limit Theorems,
which describe the distributions of sums of a great many independent and exchangeable things:
the \emph{stable laws} \citep{gnedenko1968limit}.
Roughly speaking, these say that if we add together $n$ independent copies of a random variable $X$
for which $\P\{ X > t \}$ is proportional to $t^{-\alpha}$ for some $0 < \alpha < 2$,
then -- almost regardless of the details -- the result, scaled by $n^{1/\alpha}$,
is well-approximated by a universal form, an $\alpha$-stable distribution
(and the approximation becomes exact, as $n \to \infty$).
In such ensembles, single large entries can be important:
concretely, the largest of the $n$ copies will be of order $n^{1/\alpha}$,
so that e.g., for $\alpha=1$ a positive proportion of the sum of $n$ entries will be contributed by a single one.

@@

Stable distributions are by no means new to population biology.
For instance, it is a long-standing question how often the motions of organisms
are better modeled by essentially Brownian random walks or by L\'evy processes \citep{benhamou2007animals}
and a growing literature studies the effects of such power law distributions
on spatial genetic patterns and evolutionary dynamics
\citep[e.g.,][]{paulose2019spatial,smith2023isolation}.
More directly analogous to the problem we study here, \citet{landis2012phylogenetic}
found that an $\alpha$-stable process (with $\alpha \approx 3/2$) fit the evolution of log endocranial volume
on the primate phylogeny
better than a Brownian model.
It is also important to note that
the infinitesimal model implies a Gaussian distribution for the difference between an offspring's trait and their parents' average,
\emph{not} for the entire population.
The question of when the trait distribution in an evolving population
might be modeled as a Gaussian has been discussed extensively
\citep[e.g.,][]{turelli1994genetic,dbarre2015evolution}.

The ways in which genetic variants affect an organisms' trait can be complex,
and interactions between alleles, between loci, and with the environment are often important in practice.
In fact, many traits claimed to be genetically determined
in fact only appear heritable through non-genetic means,
e.g., systematic inequalities in resource distribution partitioned by class or race.
% \citep[e.g.,][]{davenport1916hereditary}.
In this paper, we follow the literature in working with the relatively simple purely additive case
of a trait that can be decomposed into a sum across loci of the contributions of the alleles from each locus,
even omitting an independent effect of ``the environment''.
However, we emphasize that such models should not always be the default for understanding biological variation,
especially for human traits.

@@

Suppose that the focal individual's trait is
$$
    Z_M = \sum_{\ell=1}^M \eta_\ell / M^{1/\alpha},
$$
where $\eta_\ell$ is the contribution of the allele carried at the $\ell^\text{th}$ locus.
The effects of the alleles carried by the two parents
are $\eta_\ell^{[1]}$ and $\eta_\ell^{[2]}$, respectively,
and their traits are
$Z_M^{[1]} = \sum_\ell \eta_\ell^{[1]} / M^{1/\alpha}$ and
$Z_M^{[2]} = \sum_\ell \eta_\ell^{[2]} / M^{1/\alpha}$.
We assume the loci are unlinked and inheritance is fair, so
letting $X_\ell$ be a family of Bernoulli(1/2) random variables that are independent of everything else,
$$
    \eta_\ell = X_\ell \eta_\ell^{[1]} + (1 - X_\ell) \eta_\ell^{[2]} .
$$
What we will call the \emph{Mendelian sampling term}
(or occasionally \emph{Mendelian noise}), $R_M$,
is the offspring's trait minus the midparent value, $\bar Z_M$:
\begin{align*}
    \bar Z_M &= \frac{Z_M^{[1]} + Z_M^{[2]}}{2} \\
    R_M &= Z_M - \bar Z_M \\
        &= \sum_{\ell=1}^M
            \frac{
                X_\ell (\eta_\ell^{[1]} - \eta_\ell^{[2]})
                + (1 - X_\ell) (\eta_\ell^{[2]} - \eta_\ell^{[1]})
            }{ 2 } .
\end{align*}
Note in particular that
the contribution of locus $\ell$ to $R_M$ is $\pm |\eta_\ell^{[1]} - \eta_\ell^{[2]}|$.

@@

It is mathematically intriguing to speculate what effect a power law effect size distribution
for would have on the induced evolutionary dynamics.
But, is there any evidence that non-Gaussian effects are important in practice?
In this paper we focus primarily on the theory,
but first take a brief and imperfect look at what some data have to say on the topic.
To do this, we downloaded the GWAS results provided by~\citet{biobankSNPs}
for 7,221 phenotypes from 361,194 human individuals from the March 2018 release of the UK Biobank.
After filtering (see details below) we were left with 695 binary illness-related phenotypes.
Taken at face value, these provide for each phenotype a set of estimates of the additive effect
of the alternative allele at a set of single nucleotide polymorphisms (SNPs)
on the phenotype relative to the reference allele.
The phenotypes we consider were all binary and coded as 0/1,
so estimates could be thought of as additive adjustments to the risk
(so, effects are \emph{not} on a logit scale).
These estimates were obtained as the coefficient for the SNP in a simple linear model
fit to each phenotype for each SNP,
with the first 20 principal components and all combinations of age, age squared, and inferred sex as covariates;
we considered effects of those SNPs with a reported $p$-value less than $10^{-8}$.
There are a great many possible issues with these data;
however, the results seem consistent across many SNPs and phenotypes,
and we know of no reason that possible methodological artifacts
(e.g., confounding by uncorrected variables)
would specifically affect the decay of the tail of effect sizes.

Since the mendelian sampling term is a sum of terms of the form
$$
    (X_\ell - 1/2) \eta_\ell^{[1]} + (1/2 - X_\ell) \eta_\ell^{[2]},
$$
where $\eta_\ell^{[i]}$ is the effect of the allele in parent $i$ at locus $\ell$,
the appropriate central limit theorem for the sum is determined by
the tail behavior of $|(X - 1/2) \eta| = |\eta|/2$,
where $X$ is Bernoulli(1/2) and $\eta$ is the effect
of a uniformly chosen allele at a uniformly chosen locus.
So, we'd like to know
if $\P\{|X|>2t\}$ is proportional to $t^{-\alpha}$,
i.e., the value of
$$
    \lim_{t \to \infty} \frac{ \log \P\{ |X| > 2 t \} }{ \log t } = \alpha ,
$$
if the limit exists.
Let $p_\ell$ be the minor allele frequency at the $\ell^\text{th}$ locus;
since the data we have assigns effect 0 to the major allele,
$\eta_\ell = 0$ with probability $1-p_\ell$.
Therefore, if we have estimated effects for $M$ SNPs
with minor allele frequency $p_\ell$ and effect size $e_\ell$ for the minor allele,
for $1 \le \ell \le M$,
then
$$
    \P\{ |\eta| > 2t \} \approx \frac{1}{M} \sum_{\ell : |e_\ell| > 2t} p_\ell .
$$
So, to obtain a per-trait estimate of $\alpha$,
we plot $\log \frac{1}{M} \sum_{\ell : |e_\ell| > t} p_\ell$ against $\log t$
for a sequence of values of $t$,
and fit a linear model to the tail
(which we take to be all values of $t$ for which the number of SNPs with absolute effect above $t$
is between 20 and 10\% of all SNPs).
Examples are shown in Figure~\ref{fig:example_snps}.

@@

Perhaps surprisingly, there is a converse --
in at least some situations,
independence of the Mendelian sampling term implies a Gaussian distribution.
To see this,
divide the parental loci into groups according to the results of inheritance,
defining
\begin{align*}
    U_M^{[1]} &= \sum_{\ell=1}^M X_\ell \eta_\ell^{[1]}  ,
    \qquad \qquad
    V_M^{[1]} = \sum_{\ell=1}^M (1-X_\ell) \eta_\ell^{[1]} ,  \\
    U_M^{[2]} &= \sum_{\ell=1}^M (1-X_\ell) \eta_\ell^{[2]} ,
    \qquad \qquad
    V_M^{[2]} = \sum_{\ell=1}^M X_\ell \eta_\ell^{[2]}  ,
\end{align*}
so that 
$Z_M^{[1]} = U_M^{[1]} + V_M^{[1]}$ and
$Z_M^{[2]} = U_M^{[2]} + V_M^{[2]}$, while
$Z_M = U_M^{[1]} + U_M^{[2]}$ is the transmitted portion of the parental genomes,
and we can define $V_M = V_M^{[1]} + V_M^{[2]}$ to be the untransmitted portion.
Furthermore,
\begin{align*}
    \bar Z_M = \frac{Z_M + V_M}{2}
    \qquad \text{ and } \qquad
    R_M = \frac{Z_M - V_M}{ 2 } .
\end{align*}

@@

The Kac-Bernstein theorem states that if two independent random variables
have the property that their sum and difference are independent of each other,
then the random variables have a joint Gaussian distribution
(\citet{kac1939characterization,bernstein1941property},
and see \citet{kagan1973characterization} for many much more general results).
Since $R_M$ is (half) the difference and $\bar Z_M$ is (half) the sum
of $Z_M$ and $V_M$, and $Z_M$ and $V_M$ are independent,
if we assume that also $R_M$ and $\bar Z_M$ are independent,
then both are therefore jointly Gaussian.

@@

\begin{prop}\label{prop:parental_contribs}
    Suppose that $\{\eta_\ell^{[i]}\}$ are independent for all $\ell$ and $i$,
    that the distribution of $\eta_\ell^{[i]}$ only depends on $\ell$,
    and that $\{X_\ell\}$ are independent of $\{\eta_\ell^{[i]}\}$.
    If the Mendelian sampling term $R_M$ is independent of the midparent value $\bar Z_M$,
    then $R_M$ and $\bar Z_M$ are jointly Gaussian.
\end{prop}

@@

Since models can be easily set up for which trait distributions are not Gaussian,
the implication of this is that for such models, independence of the Mendelian sampling term
is not likely a good assumption.
Note that the proof does not depend on independence of the $X_\ell$:
indeed, alleles may be inherited together,
although in practice this would quickly create dependencies between nearby allelic effects.
Also note that it is \emph{not} usually true that the components --
for instance, $U_M^{[1]}$ and $U_M^{[2]}$ --
are independent of each other.

As we discuss more later,
the presence of a large effect polymorphic allele
makes $R$ and $\bar Z$ dependent:
consider the extreme case where $\eta^{[i]}_\ell$ are i.i.d.{} with variance 1 for $\ell \ge 2$,
while $\eta^{[i]}_1 = \pm 10$ with probability 1/2 each,
and $Z^{[i]} = \eta_1^{[i]} + \sum_{\ell=2}^M \eta_\ell^{[i]} / \sqrt{M}$.
Then in the limit $R$ depends strongly on parental trait values:
if $Z^{[1]} \approx Z^{[2]}$ and so $\bar Z \approx \pm 10$ then most likely
$\eta_1^{[1]} = \eta_1^{[2]}$ and so $R$ is Normal with variance 1/2;
but if $\eta_1^{[1]} \neq \eta_1^{[2]}$ (and so $\bar Z \approx 0$)
then $R$ is close to $+5$ or $-5$ with equal probability.

@@

Concretely, consider a  Moran process where an individual whose trait value is $Z$ dies at rate $\mu(Z)$
and is replaced by a the offspring of two randomly chosen individuals,
i.e., a new individual of phenotype
\begin{align*}
  \frac{Z^{1} +Z^{2}}{2} + R ,
\end{align*}
where $Z^{1}$ and $Z^{2}$ are independent draws from the empirical population distribution
and $R$ is an independent draw from the Mendelian sampling distribution
(independent of $Z^{1}$ and $Z^{2}$). 
Mutation-selection balance demands that that
the distribution of individuals removed by death
and the distribution of individuals added by birth are equal,
\ie that $(Z^{1} +Z^{2})/2 + R$ is equal in distribution
to a $\mu$-weighted choice of $Z$.
% In other words, for all bounded and continuous functions $f$,
% \begin{equation} \label{eqn:MSB_condition}
%     \E\left[ f\left(\frac{Z^{1} +Z^{2}}{2} + R \right) \right] = \frac{\E\left[ f(Z) \mu(Z) \right]}{\E[\mu(Z)]},
% \end{equation}
% where the expectation is with respect to the joint law of the population empirical distribution (the law of $Z$, $Z^{1}$ and $Z^{2}$), and $R$.
% 
% Since the set of functions $\mathcal{S} = \{e^{i \theta x} : \theta \in \R\}$ is separating (see \eg \S 3.4 in \cite{Ethier+Kurtz1986}), it suffices to show that \eqref{eqn:MSB_condition} holds for $f \in \mathcal{S}$, \ie
Taking the characteristic function of both these distributions,
deterministic mutation-selection balance is therefore equivalent to
\begin{align} \label{eqn:MSB_CF}
    \E\left[ e^{i\theta\left(\frac{Z^{1} +Z^{2}}{2} + R \right)}\right] = \frac{\E\left[ e^{i \theta Z}\mu(Z) \right]}{\E[\mu(Z)]}.
\end{align}
Now, we assumed that $Z^{1}$, $Z^{2}$, and $R$ are independent, so we can write \eqref{eqn:MSB_CF} as 
\begin{align*}
    \frac{\E\left[ e^{i\theta Z} \mu(Z) \right]}{\E[\mu(Z)]} &= \E\left[ e^{i\theta\left(\frac{Z^{1} +Z^{2}}{2} + R \right)}\right]\\
	% &= \E\left[ e^{i\frac{\theta}{2} Z^{1}}e^{i\frac{\theta}{2} Z^{2}}e^{\i\theta R}\right]\\
	&= \E\left[e^{i\frac{\theta}{2} Z^{1}}\right]\E\left[e^{i\frac{\theta}{2} Z^{2}}\right]\E\left[e^{i\theta R}\right]\\
	&= \phi\left(\frac{\theta}{2}\right)^{2} \Phi(\theta),
\end{align*}
where
$ \phi(\theta) = \mathbb{E}\left[e^{i\theta Z^{1}}\right] = \mathbb{E}\left[e^{i\theta Z^{2}}\right] $
and
$ \Phi(\theta) = \mathbb{E}\left[e^{i\theta R}\right] $
are the characteristic functions of $Z^{j}$ and $R$, respectively.  In what follows, we will explore some non-Gaussian solutions $\phi(\theta)$ to \eqref{eqn:MSB_CF}.

@@

The key observation is this:
while is meaningless to speak of variances when the Mendelian sampling term is $\alpha$-stable, if an individual's trait value is informative of the allelic effect at any given locus,
then its trait value is informative of the \emph{distribution} of an offsprings' trait value,
and hence there can be no independence between parental traits and the Mendelian sampling term
that is central to the tractability of the infinitesimal model.

@@

\begin{figure}
    \begin{center}
        \includegraphics{sims/neutral_seg_noise}
    \end{center}
    \caption{
        Segregation noise with Gaussian \textbf{(A-C)} and Cauchy \textbf{(D-F)} effect size distributions,
        from the same simulations shown in Figure~\ref{fig:trait_distrns},
        for all individuals born during the last 1,000 time units of the simulation.
        \textbf{(A,D)} Segregation noise (offspring trait minus midparent) plotted against midparent.
        \textbf{(B,E)} Histograms of the Mendelian sampling term.
        \textbf{(B,E)} Enrichment in quantiles of the Mendelian sampling term conditional on midparent value.
        To compute the latter,
        let $q_0, q_2, q_4, \ldots, q_{98}, q_{100}$ divide the distribution of the Mendelian sampling term
        into 50 bins with roughly equal numbers in each (i.e., the quantiles),
        let $p_k(a,b)$ denote the proportion of the offspring of parents with midparent value
        is in $[a, b)$ whose Mendelian sampling term is in $[q_{k-1}, q_k)$,
        and let $p_k$ be the proportion of all offspring with segregation noise in $[q_{k-1}, q_k)$
        (so, $p_k \approx 0.02$).
        Then, each line shows $p_k(a,b)/p_k$
        plotted against the midpoint of the corresponding quantile bin $(q_{k-1},q_k)$,
        for $(a,b)$ chosen to span the 5\% of midparents centered on:
        (red) the 10\% quantile,
        (black) the median, and
        (green) the 90\% quantile of midparent value.
        \label{fig:seg_noise}
    }
\end{figure}

We first consider the neutral case (i.e., with $\mu \equiv 1$).
As shown in Figure~\ref{fig:trait_distrns},
a simulation with a Cauchy effect size distribution, unsurprisingly,
has a distribution of trait values with more extreme values
and a rougher path of population median trait value over time
than does a simulation with a Gaussian effect size distribution.
Indeed, the sum of all effects along the path from the start of the simulation
to an arbitrary individual will approximate either Brownian motion (in the Gaussian case)
or a Cauchy process (in the Cauchy case).
Figure~\ref{fig:seg_noise} shows how the Mendelian sampling term
depends on midparent value.
In the Gaussian case, the Mendelian noise is Gaussian and independent of midparent value,
as shown by the horizontal lines at 1.0 in Figure~\ref{fig:seg_noise}C. 
On the other hand, in the Cauchy case we can see from either Figure~\ref{fig:seg_noise}D or F
that Mendelian noise depends strongly on midparent trait,
as we would expect in the presence of several large-effect alleles.

@@

However,
the two models look much more similar when considering only alleles of small effect.
Figure~\ref{fig:seg_noise_small} shows that midparent values
and Mendelian sampling terms computed using only mutations with effect size less than 0.2
(excluding less than 1\% of mutations)
are quite close to independent.
The result does not depend strongly on the cutoff chosen,
and are similar even if less 0.1\% of mutations are excluded.

@@

Further analysis of the infinitesimal model with other noise distributions,
as we have begun here,
could be interesting,
e.g., to see how changing the distribution of the Mendelian sampling term
affects the rate of adaptive evolution,
levels of genetic load,
the rate of fixation,
and the strength of linked selection.
However,
we have seen that the assumption of the infinitesimal model
implies a sort of ``blending inheritance''
for the large effect alleles,
and so such work must be accompanied by simulations
to see if the predictions are borne out under concrete models of genetic inheritance.
Another open question is whether there is some set of assumptions
(perhaps involving linkage, pleiotropy, and/or environmental interactions with genetic effects)
which leads to these non-Gaussian infinitesimal models.
Finally, it may well be that there is a clever way to set up a trait-only model of evolution
that is both insensitive to the underlying genetics
and does not rely on the Gaussian central limit theorem.
